---
title: "Analytic Models"
subtitle: "Assignment 2"
author: "Mina Mousavifar - 11279515 - sem311"
output:
  pdf_document:
  toc: true
  toc_depth: 2
---

<h1>
Question 1.
</h1>

<h2>
**a)**
</h2>
<p>
Based on Little's law so far we have:
$N = X(R + Z)$
$\newline$
Where N = number of users in system, X = access rate to webpage, R = mean residence time, Z = think time.
$\newline$
In addition access rate to disk is 16 $\frac{access}{second}$ and each webpage access needs 2 accesses to disk. So the access rate for a webpage is:
$X_{webpage}=16/2=8$ $\frac{access}{second}$
$\newline$
So:
$$36 = 8*(1.5+Z) \rightarrow 4.5 = 1.5 + Z \rightarrow Z = 3 seconds$$
</p>

<h2>
**b)**
</h2>
<p>
Based on Little's law and previous part: $N = X*Z$
$\newline$
Where N = average number of users thinking, X = access rate to webpage, Z = think time.
$$N = 3*8 = 24$$
</p>

***
<h1>
Question 2.
</h1>

<h2>
**a)**
</h2>

<p>
$$answer!$$
```{r, message=FALSE}

```
</p>

<h2>
**b)**
</h2>

<p>
$$answer!$$
```{r, message=FALSE}

```
</p>

***
<h1>
Question 3.
</h1>

<h2>
**a)**
</h2>

<p>
Based on Forced flow law we have: $D_{k} = \frac{B_{k}}{C}$ where D is service demand, k is each service part, B is resource busy time and C is number of requests completed.
$$D_{processor} = \frac{B_{processor}}{C}=\frac{400}{5000} = 0.08 seconds$$
$$D_{disk1} = \frac{B_{disk1}}{C}=\frac{500}{5000} = 0.1 seconds$$
$$D_{disk2} = \frac{B_{disk2}}{C}=\frac{600}{5000} = 0.12 seconds$$
So the system service demand would be:
$$D = \sum{D_{k}} = 0.08 + 0.1 + 0.12 = 0.3 seconds$$
Therefor we find that disk2 is our bottleneck with the most service demand.
</p>

<h2>
**b)**
</h2>

<p>
Processor is made 10x faster, so it's service demand has been changed: $D_{processor}=\frac{D_{processor}}{10}=\frac{0.08}{10} = 0.008 seconds$
$\newline$
and the system service demand has been changed: 
$D_{proc10x} = \sum{D_{k}} = 0.008 + 0.1 + 0.12 = 0.228 seconds$
$\newline$
Then we calculate $N^*$ which is the intersection of heavy load and light load.
$$N^*=\frac{D + Z}{D_{max}}=\frac{0.3+2}{0.12}=\frac{2.3}{0.12}=19.1\overline{6}$$
$$N_{proc10x}^*=\frac{D_{proc10x} + Z}{D_{max}}=\frac{0.228+2}{0.12}=\frac{2.228}{0.12}=18.5\overline{6}$$
For bounds on throughput we have:
$$max(D, ND_{max}-Z) \leq R(N) \leq ND \rightarrow max(0.3, 0.12N - 2) \leq R(N) \leq 0.3N$$
$$max(D_{proc10x}, ND_{max}-Z) \leq R_{proc10x}(N) \leq ND_{proc10x} \rightarrow max(0.228, 0.12N - 2) \leq R(N) \leq 0.228N$$
```{r, message=FALSE, warning=FALSE}
library(dplyr)
# calculate lower bounds for original system
Rs <- data.frame("N"=1:50, "R"=rep(0,50), stringsAsFactors = FALSE)

for(i in 1:50){
  Rs[i,2] = max(0.3, (0.12*Rs[i,1]) - 3)
}

# calculate lower bounds for 10x faster processor
Rs10x <- data.frame("N"=1:50, "R"=rep(0,50), stringsAsFactors = FALSE)

for(i in 1:50){
  Rs10x[i,2] = max(0.228, (0.12*Rs10x[i,1]) - 3)
}

# calculate percentage decrease
des <- data.frame("N"=1:50, "change"=rep(0,50), stringsAsFactors = FALSE)

for(i in 1:50){
  des[i,2] = ((abs(Rs10x[i,2]-Rs[i,2]))*100)/Rs[i,2]
}
despre <- des

# plotting
library(ggplot2)
p <- ggplot(data = des, mapping = aes(x = N, y = change)) +  
  xlim(0,50) + ylim(0, 100) + 
  ggtitle("Percentage decrease in response time for 10x processor") + 
  xlab("customers") + ylab("percentage decrease") +
  geom_point() +
  geom_line(data = des, aes(x = N, y = change, color = "10xfaster processor"))
p
```
</p>

<h2>
**c)**
</h2>

<p>
First we should calculate the new service demand time. Number of visits still stays the same and we have the same $S_{k}$ because of our identical disks. Also, because of new load balancing new $D_k$ would be the same too. So the demand would be:
$$V_1+ V_2 =  \frac{D_1}{S} + \frac{D_2}{S} \rightarrow \frac{V_1S}{S} + \frac{V_2S}{S} + \frac{V_3S}{S} = \frac{1}{S}(0.1+0.12) \rightarrow \frac{D_1S}{S} + \frac{D_2S}{S} + \frac{D_3S}{S} =\frac{1}{S}(0.22)$$
$$D_1=D_2=D_3 \rightarrow D_1(\frac{1}{S}+\frac{1}{S}+\frac{1}{S})=\frac{1}{S}(0.22) \rightarrow D_1*3 = 0.22 \rightarrow D_1=0.07\overline{3}$$
Therefore the new bounderies would be: $D = \sum{D_k} = 0.22$ and $D_{max} = 0.07\overline{3}$
$$max(D_{newdisk}, ND_{max}-Z) \leq R_{newdisk}(N) \leq ND_{newdisk} \rightarrow max(0.22, 0.07\overline{3}N - 2) \leq R(N) \leq 0.07\overline{3}N$$
</p>

```{r, message=FALSE, warning=FALSE}
# calculate lower bounds for new disk
Rsd3 <- data.frame("N"=1:50, "R"=rep(0,50), stringsAsFactors = FALSE)

for(i in 1:50){
  Rsd3[i,2] = max(0.22, (0.073*Rsd3[i,1]) - 3)
}

# calculate percentage decrease
des <- data.frame("N"=1:50, "change"=rep(0,50), stringsAsFactors = FALSE)

for(i in 1:50){
  des[i,2] = ((abs(Rsd3[i,2]-Rs[i,2]))*100)/Rs[i,2]
}

# plotting
library(ggplot2)
p <- ggplot(data = des, mapping = aes(x = N, y = change)) +  
  xlim(0,50) + ylim(0, 100) + 
  ggtitle("Percentage decrease in response time for new disk and load balancing") + 
  xlab("customers") + ylab("percentage decrease") +
  geom_point() +
  geom_line(data = des, aes(x = N, y = change, color = "adding new disk"))
p

p <- ggplot(data = des, mapping = aes(x = N, y = change)) +  
  xlim(0,50) + ylim(0, 100) + 
  ggtitle("System Upgrades") + 
  xlab("customers") + ylab("percentage decrease") +
  geom_point() +
  geom_line(data = des, aes(x = N, y = change, color = "adding new disk")) +
  geom_point(data = despre, mapping = aes(x = N, y = change)) +
  geom_line(data = despre, aes(x = N, y = change, color = "10xfaster processor"))
p
```
<p>
So we can see that the change in the processor has minimal effect in light load which disappears in the heavy load. However, the change in the disks lead to a significant improvement, especially in heavy load condition. Because our bottleneck was the disk performance.
</p>

***
<h1>
Question 4.
</h1>

<h2>
**a)**
</h2>

<p>
$$answer!$$
</p>

```{r, message=FALSE}

```

<h2>
**b)**
</h2>

<p>
$$answer!$$
</p>

```{r, message=FALSE}

```

<h2>
**c)**
</h2>

<p>
$$answer!$$
</p>

```{r, message=FALSE}

```

***
<h1>
Question 5.
</h1>

<h2>
**a)**
</h2>

<p>
$$answer!$$
</p>

```{r, message=FALSE}

```

<h2>
**b)**
</h2>

<p>
$$answer!$$
</p>

```{r, message=FALSE}

```