---
title: "Analytic Models"
subtitle: "Assignment 2"
author: "Mina Mousavifar - 11279515 - sem311"
output:
  pdf_document:
  toc: true
  toc_depth: 2
---

<h1>
Question 1.
</h1>

<h2>
**a)**
</h2>
<p>
Based on Little's law so far we have:
$N = X(R + Z)$
$\newline$
Where N = number of users in system, X = access rate to webpage, R = mean residence time, Z = think time.
$\newline$
In addition access rate to disk is 16 $\frac{access}{second}$ and each webpage access needs 2 accesses to disk. So the access rate for a webpage is:
$X_{webpage}=16/2=8$ $\frac{access}{second}$
$\newline$
So:
$$36 = 8*(1.5+Z) \rightarrow 4.5 = 1.5 + Z \rightarrow Z = 3 seconds$$
</p>

<h2>
**b)**
</h2>
<p>
Based on Little's law and previous part: $N = X*Z$
$\newline$
Where N = average number of users thinking, X = access rate to webpage, Z = think time.
$$N = 3*8 = 24$$
</p>

***
<h1>
Question 2.
</h1>

<h2>
**a)**
</h2>

<p>
$$answer!$$
```{r, message=FALSE}

```
</p>

<h2>
**b)**
</h2>

<p>
$$answer!$$
```{r, message=FALSE}

```
</p>

***
<h1>
Question 3.
</h1>

<h2>
**a)**
</h2>

<p>
$$answer!$$
</p>

<h2>
**b)**
</h2>

<p>
$$answer!$$
</p>

<h2>
**c)**
</h2>

<p>
$$answer!$$
</p>

***
<h1>
Question 4.
</h1>

<h2>
**a)**
</h2>

<p>
$$answer!$$
</p>

```{r, message=FALSE}

```

<h2>
**b)**
</h2>

<p>
$$answer!$$
</p>

```{r, message=FALSE}

```

<h2>
**c)**
</h2>

<p>
$$answer!$$
</p>

```{r, message=FALSE}

```

***
<h1>
Question 5.
</h1>

<h2>
**a)**
</h2>

<p>
$$answer!$$
</p>

```{r, message=FALSE}

```

<h2>
**b)**
</h2>

<p>
$$answer!$$
</p>

```{r, message=FALSE}

```