---
title: "Characterizing Measurement Data"
subtitle: "Assignment 1"
author: "Mina Mousavifar - 11279515 - sem311"
date: "`r Sys.time()`"
output:
  pdf_document:
  toc: true
  toc_depth: 2
---

<h1>
Question 1.
</h1>

<h2>
**a)**
</h2>
<p>
$$\lambda=5\frac{session}{minute}\rightarrow mean=\frac{1}{\lambda}=\frac{1minute}{5session}\times\frac{60 seconds}{1 minute}=12seconds$$
</p>

<h2>
**b)**
</h2>
<p>
Exponential distribution has *memoryless property*. It means that no matter how much time has elapsed since the previous arrival, the mean waiting time at time T for the next entry will still be 12 seconds.
</p>

<h2>
**c)**
</h2>
<p>
$$arrivals = period*lambda = 2minutes*5\frac{arrival}{minute} = 10 arrivals$$
</p>

<h2>
**d)**
</h2>
<p>
Because interarrival times are IID, the first arrival doesn't change the mean.
$$time = 3*60=180seconds \rightarrow arrivals=\left \lfloor{\frac{180-18}{12\frac{arrival}{seconds}}}\right \rfloor=13 + 1(first arrival) = 14$$

***
<h1>
Question 2.
</h1>

<h2>
**a)**
</h2>

<p>
*W:*
$$mean=\frac{1}{\lambda}=12\rightarrow \lambda=\frac{1}{12}=0.08\overline{3}$$
*X:*
$$mean={n}{\gamma}\rightarrow \gamma=\frac{12}{3}=4$$
*Y:*
$$mean=e^{(\mu+\frac{\sigma^2}{2})}\rightarrow e^{(1+\frac{\sigma^2}{2})}=12 \rightarrow ln(e^{(1+\frac{\sigma^2}{2})})=ln(12) \rightarrow 1+\frac{\sigma^2}{2}=ln(12) \rightarrow \frac{\sigma^2}{2}=ln(12)-1 \rightarrow \sigma^2=2ln(12)-2$$
$$\sigma=\pm(2ln(12)-2)$$
```{r, message=FALSE}
sigma=sqrt(2*log(12) - 2)
cat("sigma is: +-",sigma)
```
*Z:*
\[mean=\left\{
                \begin{array}{ll}
                  \infty \quad for \quad  \alpha \leq1\\
                  \frac{\alpha * k}{\alpha - 1} \quad for \quad  \alpha >1
                \end{array}
              \right.\]
$$\alpha=1.25 \rightarrow mean=\frac{\alpha * k}{\alpha - 1} \rightarrow 12=\frac{1.25k}{1.25-1} \rightarrow 12=\frac{1.25k}{0.25} \rightarrow 12=5k \rightarrow k = \frac{12}{5}=2.4$$
</p>

<h2>
**b)**
</h2>

<p>
*W:*
$$F(x)=P(x<12)=1-e^{-\lambda x}=1-e^{-\frac{1}{12}*12} = 1 - \frac{1}{e}$$
```{r, message=FALSE}
p=pexp(12, rate=1/12) 
cat("Probability for exponential distribution is: ",p)
```
*X:*
$$F(x)=P(x<12)=1-\sum_{n=0}^{2}\frac{e^{-{\frac{x}{\gamma}}}x^n}{\gamma^nn!}=1- \sum_{i=0}^{2}\frac{e^{-{\frac{x}{4}}}x^n}{4^nn!}$$
```{r, message=FALSE}
p=pgamma(12, shape= 3, scale= 4) 
cat("Probability for erlang distribution is: ",p)
```
*Y:*
$$F(x)=P(x<12)=\phi(\frac{(lnx) - \mu}{\sigma})$$
where $\phi$ is the cumulative distribution function of the standard normal distribution.
```{r, message=FALSE}
p=plnorm(12, meanlog = 1, sdlog = sigma) 
cat("Probability for lognormal distribution is: ",p)
```
*Z:*
$$F(x)=P(x<12)=1-{\frac{k}{x}}^\alpha$$
```{r, message=FALSE}
library(EnvStats)
p=ppareto(12, location=2.4, shape=1.25) 
cat("Probability for pareto distribution is: ",p)
```
</p>

<h2>
**c)**
</h2>

<p>
*W:*
$$1-P(x<240)=1-(1-e^{-\lambda x})=e^{-\frac{1}{12}*240} = \frac{1}{e^{20}}$$
```{r, message=FALSE}
p= 1 - pexp(240, rate=1/12) 
cat("Probability for exponential distribution is: ",p)
```
*X:*
$$1-P(x<240)=1-(1-\sum_{n=0}^{2}\frac{e^{-{\frac{x}{\gamma}}}x^n}{\gamma^nn!})= \sum_{i=0}^{2}\frac{e^{-{\frac{x}{4}}}x^n}{4^nn!}$$
```{r, message=FALSE}
p= 1 - pgamma(240, shape= 3, scale= 4) 
cat("Probability for erlang distribution is: ",p)
```
*Y:*
$$1-P(x<240)=1-\phi(\frac{(lnx) - \mu}{\sigma})$$
where $\phi$ is the cumulative distribution function of the standard normal distribution.
```{r, message=FALSE}
p= 1 - plnorm(240, meanlog = 1, sdlog = sigma) 
cat("Probability for lognormal distribution is: ",p)
```
*Z:*
$$1-P(x<240)=1-(1-{\frac{k}{x}}^\alpha)=\frac{k}{x}^\alpha$$
```{r, message=FALSE}
library(EnvStats)
p= 1 - ppareto(240, location=2.4, shape=1.25) 
cat("Probability for pareto distribution is: ",p)
```


***
<h1>
Question 3.
</h1>

<p>
$$P_R(n)=\frac{\frac{1}{n^\alpha}}{\sum_{m=1}^{k}\frac{1}{m^\alpha}} \rightarrow i^{-\alpha}<\frac{1}{day}=\frac{1}{60*60*24 seconds}=\frac{1}{86400} \rightarrow i^\alpha > 86400 \rightarrow i^{1.25}>86400$$
</p>

```{r, message=FALSE}
library(dplyr)

powers <- data.frame("i"=0:10000000, "pow"=rep(0,10000001), stringsAsFactors = FALSE)
powers$pow <- powers$i^1.25
is <- powers %>% filter(pow > 86400)
head(is, 1)
cat("i: ", is$i[1])

powers$pow <- powers$i^0.85
is <- powers %>% filter(pow > 86400)
head(is, 1)
cat("i: ", is$i[1])
```


***
<h1>
Question 4.
</h1>

<p>

</p>

```{r, message=FALSE}

```


***
<h1>
Question 5.
</h1>

<h2>
**a)**
</h2>
<p>

</p>

<h2>
**b)**
</h2>
<p>

</p>

<h2>
**c)**
</h2>
<p>

</p>

```{r, message=FALSE}

```


