---
title: "Characterizing Measurement Data"
subtitle: "Assignment 1"
author: "Mina Mousavifar - 11279515 - sem311"
date: "`r Sys.time()`"
output:
  pdf_document:
  toc: true
  toc_depth: 2
---

<h1>
Question 1.
</h1>

<h2>
a)
</h2>
<p>
$$\lambda=5\frac{session}{minute}\rightarrow mean=\frac{1}{\lambda}=\frac{1minute}{5session}\times\frac{60 seconds}{1 minute}=12seconds$$
</p>

<h2>
b)
</h2>
<p>
Exponential distribution has *memoryless property*. It means that no matter how much time has elapsed since the previous arrival, the mean waiting time at time T for the next entry will still be 12 seconds.
</p>

<h2>
c)
</h2>
<p>
$$arrivals = period*lambda = 2minutes*5\frac{arrival}{minute} = 10 arrivals$$
</p>

<h2>
d)
</h2>
<p>
Because interarrival times are IID, the first arrival doesn't change the mean.
$$time = 3*60=180seconds \rightarrow arrivals=\left \lfloor{\frac{180-18}{12\frac{arrival}{seconds}}}\right \rfloor=13 + 1(first arrival) = 14$$
$$\left \lceil{x}\right \rceil $$
</p>
```{r, message=FALSE}
  
```

***
<h1>
Question 2.
</h1>

<h2>
a)
</h2>

<p>
W:
$$mean=\frac{1}{\lambda}=12\rightarrow \lambda=\frac{1}{12}=0.08\overline{3}$$
X:
$$mean={n}{\gamma}\rightarrow \gamma=\frac{12}{3}=4$$
Y:
$$mean=e^{(\mu+\frac{\sigma^2}{2})}\rightarrow e^{(1+\frac{\sigma^2}{2})}=12 \rightarrow ln(e^{(1+\frac{\sigma^2}{2})})=ln(12) \rightarrow 1+\frac{\sigma^2}{2}=ln(12) \rightarrow \frac{\sigma^2}{2}=ln(12)-1 \rightarrow \sigma^2=2ln(12)-2$$
$$\sigma=\pm(2ln(12)-2)$$
```{r, message=FALSE}
sigma=2*log2(12) - 2
cat("sigma is: +-",sigma)
```
Z:
\[mean=\left\{
                \begin{array}{ll}
                  \infty \quad for \quad  \alpha \leq1\\
                  \frac{\alpha * k}{\alpha - 1} \quad for \quad  \alpha >1
                \end{array}
              \right.\]
$$\alpha=1.25 \rightarrow mean=\frac{\alpha * k}{\alpha - 1} \rightarrow 12=\frac{1.25k}{1.25-1} \rightarrow 12=\frac{1.25k}{0.25} \rightarrow 12=5k \rightarrow k = \frac{12}{5}=2.4$$
</p>

<h2>
b)
</h2>

<p>
W:
$$P(x<12)=1-e^{-\lambda x}=1-e^{-\frac{1}{12}*12} = 1 - \frac{1}{e}$$
```{r, message=FALSE}
p=pexp(12, rate=1/12) 
cat("Probability for exponential distribution is: ",p)
```
X:

Y:

Z:

</p>

<h2>
c)
</h2>

<p>
W:

X:

Y:

Z:

</p>
```{r, message=FALSE}

```


***
<h1>
Question 3.
</h1>

<p>

</p>

```{r, message=FALSE}

```


***
<h1>
Question 4.
</h1>

<p>

</p>

```{r, message=FALSE}

```


***
<h1>
Question 5.
</h1>

<h2>
a)
</h2>
<p>

</p>

<h2>
b)
</h2>
<p>

</p>

<h2>
c)
</h2>
<p>

</p>

```{r, message=FALSE}

```


